HO unification from object language to meta language

### 2024-02-26 paper/main.tex
tex update

\begin{elpicode}
decision (all A x\ Pm x) :- link Pm A P, finite A,   ~\customlabel{clause:r3a}{(r3a)}~
  pi x\ decision (app[P, x]).
\end{elpicode}

@@

\paragraph{Term unification: \Uo vs. \Ue} 
The last but not least important relation we should take care of before
presenting our full algorithm aiming to unify terms of the OL in the ML and
provide the substitution produced in the ML to the OL, is term unification. This
procedure is a more powerful version of the equal predicate, since unification
checks if two terms can be equal be assigning unification variables. In our
representation, variable assignment (or refinement) is performed by modifying
the corresponding substitution mapping. We will not give an implementation of
\Uo in the world of the OL, since we are giving an implementation of it using
our algorithm, see \cref{XX}.

On the other hand, we can

>>>>

\begin{elpicode}
decision (all A x\ Pm x) :- link Pm P A, finite A,   ~\customlabel{clause:r3a}{(r3a)}~
  pi x\ decision (app[P, x]).
\end{elpicode}

@@

\paragraph{Term unification: \Uo vs. \Ue} 
The last but not least important relation we should take care of before
presenting our full algorithm aiming to unify terms of the OL in the ML and
provide the substitution produced in the ML to the OL, is term unification. This
procedure is a more powerful version of the equal predicate, since unification
checks if two terms can be equal be assigning unification variables. In our
representation, variable assignment (or refinement) is performed by modifying
the corresponding substitution mapping. We will not give an implementation of
\Uo, since we are giving an implementation of it using
our algorithm, see \cref{XX}.

\begin{figure}
  \begin{elpicode}
    type unif tm -> tm -> ho_subst -> ho_subst ->  o.
    % Base cases
    unif (app L1) (app L2) S S1 :- fold2 unif L1 L2 S S1.
    unif (lam F1) (lam F2) S S1 :- 
      pi x\ (pi S\ unif x x S S) => unif (F1 x) (F2 x) S S1.
    unif (con X) (con X) S S.
    % Assigned variables
    unif (hov N Args) T S S1 :- 
      assigned-deref N S F, beta-abs F Args T1, unif T1 T S S1.
    unif T (hov N Args) S S1 :- 
      assigned-deref N S F, beta-abs F Args T1, unif T T1 S S1.
    % Non-assigned variabless
    unif (hov N Args1) (hov M Args2) S S3 :-  
      pattern-fragment Args1, pattern-fragment Args2, 
      new S W S1, prune W Args1 Args2 Ass, 
      assign N S1 Ass S2, assign M S2 Ass S3.
    unif (hov N Args) T S S1 :- pattern-fragment Args,
      not (T = hov _ _), not_occ N S T,  
      bind Args T T1, assign N S T1 S1.
    unif T (hov N Args) S S1 :- pattern-fragment Args,
      not (T = hov _ _), not_occ N S T, 
       bind Args T T1, assign N S T1 S1.
  \end{elpicode}
  \caption{The \Ue predicate}
  \label{code:ue-pred}
  \Description[code:unif-pred]{code:unif-pred}
\end{figure}

On the other hand, unification in the ML needs to be defined. In
\cref{code:ue-pred}, we give an implementation of \Ue. This predicate has four
arguments, the two terms $t_1$ and $t_2$ to unify, the old substitution map
$\rho_1$, and the new substitution map $\rho_2$, with the invariant $\rho_1
\subseteq \rho_2$. The first three rules unify terms with same rigid heads, and
call the unification relation on the sub-terms. If $t_1$ (resp. $t_2$) is an
assigned variables, $t_1$ is dereferenced to $t_1'$ (resp. $t_2'$) and the
unification is called between $t_1'$ and $t_2$ (resp. $t_1$ and $t_2'$). If both
terms are unification variables, we test that their arguments are in the pattern
fragment, we allocate a new variable $w$ in $\rho_1$ such that $w$ is the
pruning of the arguments of $t_1$ and $t_2$, we assign both $t_1$ and $t_2$ to
$w$ and return the new mapping $\rho_2$ containing all the new variable
assignment. Finally, if only one of the two terms is an unification variable
$v$, after having verified that $v$ does not occur in the other term $t$, we
bind $v$ to $t$ and return the new substitution mapping.
